Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-07-24 3.technique.tex
update

Given an alias graph, whether a variable and a lock exist in the same data 
structure can be determined by finding a common ancestor. If they are in the 
same data structure, the lock is likely to protect the variable when it is 
accessed. Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
get the field of the accessed variable and the field of the protecting lock in 
the form of access path, if they exist in the same data structure. Given a node 
of an accessed variable and a node of a lock variable, the analysis first gets 
the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
does not exist, the analysis returns NULL (Lines 2-3). Otherwise, the analysis 
gets the access path for the node of the accessed variable and the node of the 
lock variable from the common ancestor (Lines 5-7).

@@

\PP{S1: Key-field Extraction.} The OS Kernel has a large code base with 
numerous variables. However, only a small part of variables should be protected 
by a specific lock, and thus collecting all variable accesses can introduce 
much unnecessary overhead. Generally, only a few field accesses can miss 
necessary protecting lock, due to carelessness of developers. Based on this 
insight, our analysis first extracts key fields that may need to be protected 
by a specific lock, by performing a lock-set analysis to find whether a given 
variable is accessed after acquiring a lock in the same data structure as the 
accessed field in any code path.

@@

Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to collect key 
fields that may be protected by a specific lock when are accessed, based on 
lock-set analysis and alias graph. The analysis start analyzing

For each function without a caller function, the analysis

>>>>

Given an alias graph, whether a variable and a lock exist in the same data 
structure can be determined by finding a common ancestor. If they are in the 
same data structure, the lock is likely to protect the variable when it is 
accessed. Figure~\ref{fig_pseudocode_get_access} shows the pseudocode to 
get the field of the accessed variable and the field of the protecting lock in 
the form of access path, if they exist in the same data structure. Given a node 
of an accessed variable and a node of a lock variable, the analysis first gets 
the common ancestor of the two nodes (Line 1). And then, if the common ancestor 
does not exist, the analysis returns a NULL pair (Lines 2-3). Otherwise, the 
analysis gets the access path for the node of the accessed variable and the 
node of the lock variable from the common ancestor (Lines 5-7).

@@

\PP{S1: Key-field extraction.} The OS Kernel has a large code base with 
numerous variables. However, only a small part of variables should be protected 
by a specific lock, and thus collecting all variable accesses can introduce 
much unnecessary overhead. Generally, only a few field accesses can miss 
necessary protecting lock, due to carelessness of developers. Based on this 
insight, our analysis first extracts key fields that may need to be protected 
by a specific lock, by performing a lock-set analysis to find whether a given 
variable is accessed after acquiring a lock in the same data structure as the 
accessed field in any code path.

@@

Figure~\ref{fig_pseudocode_field_extract} shows the pseudocode to collect key 
fields that may be protected by a specific lock, based on lock-set analysis and 
the alias graph. The analysis starts from each function without a caller 
function (Lines 36-40) and performs a path-based analysis~\cite{Li:ASPLOS22} 
(Lines 26-31). For each instruction in the code path, the analysis first update 
the alias graph according to the instruction with the four operations (MOVE, 
STORE, LOAD and GEP) (Line 1), and then performs a lock-set analysis (Line 2) 
to get all locks acquired. After update the alias graph and the lock set, the 
analysis gets the node of the operand from the new alias graph (Lines 3-4), and 
then update the set of key fields, according to the type of instructions (Lines 
5-24). Specifically, if the instruction is a write or a read, for each node of 
lock variable in the lock set, the analysis uses the alias graph to extract the 
protected data structure field (Lines 9-11). If the field (in the form of 
access path) is not NULL, it is inserted into the set of key fields. As for a 
function call instruction, the analysis first analyzes the called function with 
the alias graph and lock set of the caller function (Lines 18-21), and then 
merges the returned key fields to the caller function (Line 22).

\PP{S2: Field-access collection.}
